Type Refinements for the Working Class

### 2015-11-27 paper.tex
Fix chktex warnings
Restore "Constructions on Refinements"

\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
\begin{quote}
  To know $\FRMemEq{\Upsilon}{\Psi,x:\chi}{N_0}{N_1}{\phi}$ is to know, for any renaming
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
  and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
  $\FRMemEq{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{N_0\rho}}{\Subst{M_1}{x}{N_1\rho}}{\Subst{M_0}{x}{\phi\rho}}$.
  In other words:
  %
  \[
    \infer{
      \FRMemEq{\Upsilon}{\Psi,x:\psi}{N_0}{N_1}{\phi}
    }{
      \begin{array}{l}
        \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}
        .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
        \Rightarrow
        \left(%
          \FRMemEq{\Upsilon'}{\Psi\rho}{%
            \Subst{M_0}{x}{N_0\rho}
          }{%
            \Subst{M_1}{x}{N_1\rho}%
          }{\Subst{M_0}{x}{\phi\rho}
          }%
        \right)
      \end{array}
    }
  \]
\end{quote}
%%
%% \section{Constructions on Refinements}
%%
%% Refinements can be arranged in a partial order via the
%% $\IsSubrefinement{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
%% at any sort $\tau$.
%%
%% We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which simply
%% contains no elements. Next, the top refinement $\GloballyRefines{\TyTop{\tau}}{\tau}$
%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
%% \begin{gather*}
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }\qquad
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \end{gather*}
%%
%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
%% dually:
%% \[
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi} &
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \]
%%
%% It is trivial to verify that these constructions on refinements give rise to a
%% lattice.
%%
%% \section{Multi-Sorted Computational Type Theory}
%%
%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
%% you specify how to form its equal canonical inhabitants, or values. Then, the
%% inhabitants of the types are explained uniformly by extending this relation
%% over an operational semantics such that two terms are equal members of a
%% type just when they compute to equal values of that type.
%%
%% More generally, one may wish to consider types which include computations in
%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
%% is a type, which would include all the members of $\psi$ as well as a divergent
%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
%% closed terms of sort $\tau$.
%%
%% In order to ensure that such types are sensible, we must formulate a notion of
%% computational equivalence, and then say that a refinement is a \emph{type} just
%% when it respects computational equivalence; then, we are able to define types
%% as PERs on \emph{programs} in addition to values.

>>>>

\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.
\begin{quote}
  To know $\FRMemEq{\Upsilon}{\Psi,x:\chi}{N_0}{N_1}{\phi}$ is to know, for any renaming
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
  and closed terms $\Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}$, that from
  $\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}$, you can conclude
  $\FRMemEq{\Upsilon'}{\Psi\rho}{\Subst{M_0}{x}{N_0\rho}}{\Subst{M_1}{x}{N_1\rho}}{\Subst{M_0}{x}{\phi\rho}}$.
  In other words:
  %
  \[
    \infer{
      \FRMemEq{\Upsilon}{\Psi,x:\psi}{N_0}{N_1}{\phi}
    }{
      \begin{array}{l}
        \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}
        .\ \forall \Member{M_0,M_1}{\Exprs_\sigma(\Upsilon')}.\\
        \left(\FRMemEq{\Upsilon'}{\Psi\rho}{M_0}{M_1}{\chi\rho}\right)
        \Rightarrow
        \left(%
          \FRMemEq{\Upsilon'}{\Psi\rho}{%
            \Subst{M_0}{x}{N_0\rho}
          }{%
            \Subst{M_1}{x}{N_1\rho}%
          }{\Subst{M_0}{x}{\phi\rho}
          }%
        \right)
      \end{array}
    }
  \]
\end{quote}

\section{Constructions on Refinements}

Refinements can be arranged in a partial order via the
$\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
at any sort $\tau$.

We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which
simply contains no elements. Next, the top refinement
$\GloballyRefines{\TyTop{\tau}}{\tau}$ identifies all terms
$\Member{M,N}{\Exprs_\tau(\Upsilon)}$.  The join
$\Refines{\Upsilon}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined as
follows:

\begin{gather*}
  \infer{
    \RMemEq{\Upsilon}{M}{N}{\phi\TyJoin{\tau}\psi}
  }{
    \RMemEq{\Upsilon}{M}{N}{\phi}
  }\qquad
  \infer{
    \RMemEq{\Upsilon}{M}{N}{\phi\TyJoin{\tau}\psi}
  }{
    \RMemEq{\Upsilon}{M}{N}{\psi}
  }
\end{gather*}

Finally the meet $\Refines{\Upsilon}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$ is defined
dually:
\[
  \infer{
    \RMemEq{\Upsilon}{M}{N}{\phi\TyMeet{\tau}\psi}
  }{
    \RMemEq{\Upsilon}{M}{N}{\phi} &
    \RMemEq{\Upsilon}{M}{N}{\psi}
  }
\]

It is trivial to verify that these constructions on refinements give rise to a
lattice.

\section{Multi-Sorted Computational Type Theory}

Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
you specify how to form its equal canonical inhabitants, or values. Then, the
inhabitants of the types are explained uniformly by extending this relation
over an operational semantics such that two terms are equal members of a
type just when they compute to equal values of that type.

More generally, one may wish to consider types which include computations in
addition to values; an example of such a type would be $\psi_\bot$ when $\psi$
is a type, which would include all the members of $\psi$ as well as a divergent
computation. Another example is the type $\TyTop{\tau}$ which identifies all
closed terms of sort $\tau$.

In order to ensure that such types are sensible, we must formulate a notion of
computational equivalence, and then say that a refinement is a \emph{type} just
when it respects computational equivalence; then, we are able to define types
as PERs on \emph{programs} in addition to values.
